Compositional equivalences based on Open pNets\thanks{This work was partially funded by the Associated Team FM4CPS between INRIA and ECNU, Shanghai}

### 2019-04-23 2016-Weak-FH-Bisim/WeakBisim.tex
update todos
preuve en cours; mardi matin

@@

- on remet l exemple de enable, etendu pour weak (Eric)

- reste nos 5 thm a prouver: transitivite, decomposition/ recomposition, congruence x 2

@@

- target SIAvocs ou SCP ou JLamp

@@

\begin{lemma}[Open transition composition]\label{lem-compose} 
	Suppose $j_0\in J$ and:\\[-2ex]
\begin{mathpar}
%\[
P\models{\openrule
	{
		\beta_j^{j\in J}, 
		\Pred,  
		\Post}
	{\ostate{s_i^{i\in L}} \OTarrow {\alpha}
		\ostate{s_i'^{\, i\in L}}}
}%\]
\text{~~and~~}
%\[
Q\models{\openrule
	{
		\beta_j^{j\in J_Q},
		 \Pred\,',  
		\Post\,'}
	{\ostate{s_i^{i\in L_Q}} \OTarrow {\alpha_Q}
		\ostate{s_i'^{\, i\in L_Q}}}
}%\]
\end{mathpar}
Then, we have\\[-2ex]
	\[ P[Q]_{j_0}  
	\models
	{\openrule
		{
			\beta_j^{(j\in J\setminus\{j_0\}) \uplus J_Q}, 
			\Pred\land\Pred\,'\land \alpha_Q=\beta_{j_0},  
			\Post\uplus \Post\,'}
		{\ostate{s_i^{i\in L\uplus L_Q}} \OTarrow {\alpha}
			\ostate{s_i'^{\, i\in L\uplus L_Q}}}
	}
	\]
\end{lemma}
\TODO{important comment added:}
Note that this does not mean that any two pNets can be composed and produce an open 
transition. Indeed, the predicate $\Pred\land\Pred\,'\land \alpha_Q=\beta_{j_0}$ is not 
be satisfiable if the action of $\alpha_Q$ cannot be matched with $\beta_{j_0}$.
Note also that $\beta_{j_0}$ is now only used as an intermediate term inside formulas: it 
does not appear neither as global action nor as an action of a hole.
\begin{small}

@@

\begin{proof} Suppose we have:
\begin{mathpar}
%\[
P\models{\openrule
	{
		\gamma_j^{j\in J}, 
		\Pred,  
		\Post}
	{\ostate{s_i^{i\in L}} \OTWeakarrow {\alpha}
		\ostate{s_i'^{\, i\in L}}}
}
\end{mathpar}
By Lemma \ref{lem-rel-OT-WOT} this implies the following: \\			
$\forall p\!\in\![0..m_1].\openrule
    {\set{\beta^1_p},\Pred^1_p,\Post^1_p   }
         {\ostate{{(s^1_{pi})}^{i\in L}} \OTarrow {\tau} \ostate{{(s^1_{(p+1)i})}^{i\in L}}}\!\in\!\mathcal{T} \wedge
\openrule
         {\set{\beta^2},\Pred^2,\Post^2 }
         {\ostate{({s_{0i}^2})^{i\in L}} \OTarrow {\alpha} \ostate{({s_{1i}^2})^{i\in L}}}\! \in\! \mathcal{T}
\wedge\\
\forall p \in [0..m_2].\openrule
         {
           \set{\beta^3_p},\Pred^3_p, \Post^3_p}
         { \ostate{(s^3_{pi})^{i\in L}} \OTarrow {\tau} \ostate{(s^3_{(p+1)i})^{i\in L}}} \in \mathcal{T}$ 
\\
where for all $i\in L$:\\
$s_i=s^1_{0 i} \wedge s^1_{(m_1+1) i}=s^2_{0i} \wedge  s^2_{1i}= s^3_{0 i} \wedge s^3_{(m_2+1) i}=s'_i$\\

@@

Moreover, we have for all $p\in[0..n]$:
\begin{mathpar}
Q\models{\openrule
	{
		\gamma_{p j}^{j\in J_p},
		 \Pred_p,  
		\Post_p}
	{\ostate{s_{p i}^{i\in L_Q}} \OTWeakarrow {\alpha_p}
		\ostate{s_{(p+1) i}^{\, i\in L_Q}}}
}		
\end{mathpar}
In the same way,  by Lemma \ref{lem-rel-OT-WOT} we can decompose $\WOT_p$ into  a series of $m'+n'$ tau open transitions and a single observable open transition:
\[Q\models{\openrule
			{
				(\beta^2_j)^{j\in J\cap\Holes(Q)}, {{\Pred'^{2}}},  
				{{\Post'^{2}}}}
				{\ostate{s_{i}^{i\in L_Q}} \OTarrow {\alpha_p}
				\ostate{s_{i}'^{i\in L_Q}}}
				}%\]
\]

We proceed by iteration on the list of   $\symb{OT}_p$ of $P$ identified above with an additional variable $t$, OTs are of the form: 
$P\!\models		
\openrule
    {
       {(\beta^i_{p j})^{j\in J}},\Pred_p,\Post_p   }
         {\ostate{s^{i\in L}_{pi}} \OTarrow {\alpha} \ostate{s^{i\in L}_{(p+1)i}}}$ 
At the beginning $Q$ is at state $\ostate{s_{0}}$ and $t=0$. 
For each step, we have two cases:
\begin{itemize}
\item If $j_0\in J$ then Q moves by one open transition. By the hypothesis above $\beta^i_{j_0}\neq \tau$. Additionally, we have
$\displaystyle{\gamma_j^{j\in J}=\mybigdotcup_{p=0}^{m_1} \vis{\set{\beta^1_p}} \dotcup  \vis{\set{\beta^2}} \dotcup \mybigdotcup_{p=0}^{m_2} \vis{\set{\beta^3_p}}}$ and\\
 $\gamma_{j_0}=[\beta_0..\beta_n]$, thus
  $\beta^i_{j_0}=\beta_t$ with  for some $t$ such that $t$ has value $0$ at beginning and that is incremented whenever this condition is satisfied.

On the other hand, by hypothesis we have:
\[Q\models{\openrule
	{
		\gamma_{t j}^{j\in J_t},
		 \Pred_t,  
		\Post_t}
	{ \ostate{s_{t}}\OTWeakarrow {\alpha_t}
		\ostate{s_{{(t+1)}}}}
}
\]
By  Lemma \ref{lem-rel-OT-WOT} we can decompose $\WOT_t$ into  a series of $k+k'$ tau open transitions and the observable open transition:

\[Q\models{\openrule
	{
		\gamma_{t j}^{j\in J_t},
		 \Pred_t,  
		\Post_t}
	{\ostate{s^2_{t}}\OTarrow {\alpha_t}
		\ostate{s^3_{t}}}
}
\]
\begin{enumerate}

\item For the first $k$ open tau transitions, by Definition \ref{def:Non-ObsTau} $P$ can make a tau open transition and the hole makes also tau action. So by Lemma \ref{lem-compose} we obtain $k$ open transitions in the form: 
\[P[Q]_{j_0}\!\models		
\openrule
    {
       \beta_p^{(p\in J\setminus\{j_0\}) \uplus J_Q},{\Pred_p},{\Post_p}   }
         {\ostate{s^{i\in L \uplus L_Q}_{pi}} \OTarrow {\tau} \ostate{s^{i\in L\uplus L_Q }_{(p+1)i}}}\]

\item We proceed in the same way for $k'$ last weak open transitions, and we obtain $k'$ open tau transitions.\\
\item For the observable open transition. Because of the definition of $\Pred'$ we can suppose that $\alpha_t=\beta_{j_0}$ \TODO{because else the original WOT does not trigger}. By Lemma \ref{lem-compose} we obtain:\\ 
\[P[Q]_{j_0}\!\models		
\openrule
    {
       \beta_p^{(p\in J\setminus\{j_0\}) \uplus J_Q},{\Pred_p}\!\!'',{\Post_p}\!\!''   }
         {\ostate{s^{i\in L \uplus L_Q}_{pi}} \OTarrow {\alpha} \ostate{s^{i\in L\uplus L_Q }_{(p+1)i}}}\] \\
 such that  ${\Pred_p}\!\!'' \iff {\Pred_p}
		\land {\Pred_p}\!\!'\land \alpha_t=\beta_{j_0}$, ${\Post_p}\!\!''={\Post_p}\uplus 
		{\Post_p}\!\!'$\\ 
\end{enumerate}

>>>>

TODOS: AVRIL 2019

- exemple (Eric: failure buffer)
- biblio (rabea)
- inuitions de preuves et intuition de bisim de cloture, etc. (Ludo)

===> dealine 14 Juillet

@@

@@

- target: JLamp

@@

\begin{lemma}[Open transition composition]\label{lem-compose}

\end{lemma}
\TODO{important comment added:}
Note that this does not mean that any two pNets can be composed and produce an open 
transition. Indeed, the predicate $\Pred\land\Pred\,'\land \alpha_Q=\beta_{j_0}$ is not 
be satisfiable if the action of $\alpha_Q$ cannot be matched with $\beta_{j_0}$.
Note also that $\beta_{j_0}$ is now only used as an intermediate term inside formulas: it 
does not appear neither as global action nor as an action of a hole.
\begin{small}

@@

\begin{proof} Suppose we have:
\begin{mathpar}
%\[
P\models{\openrule
	{
		\gamma_j^{j\in J}, 
		\Pred,  
		\Post}
	{\ostate{s_i^{i\in L}} \OTWeakarrow {\alpha}
		\ostate{s_i'^{\, i\in L}}}
}
\end{mathpar}
By Lemma \ref{lem-rel-OT-WOT} this implies the following: \\			
$\forall p\!\in\![0..m_1].P\models\openrule
    {\set{\beta^1_p},\Pred^1_p,\Post^1_p   }
         {\ostate{{(s^1_{pi})}^{i\in L}} \OTarrow {\tau} \ostate{{(s^1_{(p+1)i})}^{i\in L}}} \wedge
P\models\openrule
         {\set{\beta^2},\Pred^2,\Post^2 }
         {\ostate{({s_{0i}^2})^{i\in L}} \OTarrow {\alpha} \ostate{({s_{1i}^2})^{i\in L}}}
\wedge\\
\forall p \in [0..m_2].P\models\openrule
         {
           \set{\beta^3_p},\Pred^3_p, \Post^3_p}
         { \ostate{(s^3_{pi})^{i\in L}} \OTarrow {\tau} \ostate{(s^3_{(p+1)i})^{i\in L}}}$ 
\\
where for all $i\in L$:\\
$s_i=s^1_{0 i} \wedge s^1_{(m_1+1) i}=s^2_{0i} \wedge  s^2_{1i}= s^3_{0 i} \wedge s^3_{(m_2+1) i}=s'_i$\\

@@

Moreover, we have for all $p\in[0..n]$:
\begin{mathpar}
Q\models{\openrule
	{
		\gamma_{p j}^{j\in J_p},
		 \Pred_p,  
		\Post_p}
	{\ostate{s_{p i}^{i\in L_Q}} \OTWeakarrow {\alpha_p}
		\ostate{s_{(p+1) i}^{\, i\in L_Q}}}
}		
\end{mathpar}
%In the same way,  by Lemma \ref{lem-rel-OT-WOT} we can decompose $\WOT_p$ into  a series of $m'+n'$ tau open transitions and a single observable open transition:
%\[Q\models{\openrule
%			{
%				(\beta^2_j)^{j\in J\cap\Holes(Q)}, {{\Pred'^{2}}},  
%				{{\Post'^{2}}}}
%				{\ostate{s_{p i}^{i\in L_Q}} \OTarrow {\alpha_p}
%				\ostate{s_{(p+1) i}^{i\in L_Q}}}
%				}%\]
%\]

We proceed by iteration on the list of  OTs of $P$ identified above, these OTs are of the form:

$P\!\models		
\openrule
    {
       {(\beta^i_{p j})^{j\in J}},\Pred_p,\Post_p   }
         {\ostate{s^{i\in L}_{pi}} \OTarrow {\alpha} \ostate{s^{i\in L}_{(p+1)i}}}$

We use an additional variable $t$, in the iteration, it will be used as an index from $0$ to $n$, i.e. inside the OTs of $Q$.
At the beginning $t=0$, i.e. $Q$ is at state $\ostate{s_{0}}$. \TODO{on peut remplacer $s_t$ par $s_{t i}^{i\in L_Q}$}
For each step of the iteration, we have two cases:
\begin{itemize}
\item If $j_0\in J$ then Q moves by one open transition. By the hypothesis above $\beta^i_{p j_0}\neq \tau$. Additionally, we have 
$\displaystyle{\gamma_j^{j\in J}=\mybigdotcup_{p=0}^{m_1} \vis{\set{\beta^1_p}} \dotcup  \vis{\set{\beta^2}} \dotcup \mybigdotcup_{p=0}^{m_2} \vis{\set{\beta^3_p}}}$ and\\
 $\gamma_{j_0}=[\beta_0..\beta_n]$, thus
  $\beta^i_{p j_0}=\beta_t$ with  for some $t$ such that $t$ has value $0$ at beginning and that is incremented whenever this condition is satisfied.

On the other hand, by hypothesis we have \TODO{on peut remplacer $s_t$ par $s_{t i}^{i\in L_Q}$}:
\[Q\models{\openrule
	{
		\gamma_{t j}^{j\in J_t},
		 \Pred_t,  
		\Post_t}
	{ \ostate{s_{t}}\OTWeakarrow {\alpha_t}
		\ostate{s_{{(t+1)}}}}
}
\]
By  Lemma \ref{lem-rel-OT-WOT} we can decompose this $\WOT$ into  a series of $k+1$ and $k'+1$ tau open transitions and an observable open transition:

\[\forall h\!\in\![0..k].Q\models\openrule
    {\set{\beta^1_h},\Pred^1_h,\Post^1_h   }
         {\ostate{{(s^1_{h})}} \OTarrow {\tau} \ostate{{(s^1_{(h+1)})}}}
\]

\[Q\models{\openrule
	{
		\set{\beta^2},
		 \Pred^2,  
		\Post^2}
	{\ostate{s^2_{0}}\OTarrow {\alpha_t}
		\ostate{s^2_{1}}}
}
\]

\[
\forall h \in [0..k'].Q\models\openrule
         {
           \set{\beta^3_h},\Pred^3_h, \Post^3_h}
         { \ostate{(s^3_h)} \OTarrow {\tau} \ostate{(s^3_{(h+1)})}} \]

such that 
$s_t=s^1_{0} \wedge s^1_{(k+1) i}=s^2_{0} \wedge  s^2_{1}= s^3_{0} \wedge s^3_{(k'+1) i}=s_{t+1}$\\

$\displaystyle{\gamma_{t j}^{j\in J_t}=\mybigdotcup_{h=0}^{k} \vis{\set{\beta^1_h}} \dotcup  \vis{\set{\beta^2}} \dotcup \mybigdotcup_{h=0}^{k'} \vis{\set{\beta^3_h}}}\\
\Pred_t=\bigwedge_{h=0}^{k}\!\big(\Pred^1_h\bigotimes_{j=0}^{h-1}\subst{\Post_j^1}\big)\land\Pred^2\bigotimes_{h=0}^{k}\subst{\Post_h^1}\land \bigwedge_{h=0}^{k'}\!\Pred_h^3\big(\bigotimes_{j=0}^{h-1}\subst{\Post^3_j}\subst{\Post^2}  \bigotimes_{h=0}^{k}\subst{\Post_h^1}\big)\\
\Post=\bigotimes_{h=0}^{k'}\Post^3_h\subst{\Post^2}\bigotimes_{j=0}^{k}\subst{\Post^1_j}
$

\item For the first $k$ open tau transitions, by Definition \ref{def:Non-ObsTau} $P$ can necessarily make a tau open transition if the hole indexed $j_0$ makes a tau action. So by Lemma \ref{lem-compose} we obtain $k$ open transitions in the form: \TODO{remarquer la notation compliquee ou trouver une meilleure notation mais pas trop lourde}
\[P[Q]_{j_0}\!\models		
\openrule
    {
       \set{\beta^1_h},{\Pred^1_h},{\Post^1_h}   }
         {\ostate{s^1_h\uplus s^{i\in L}_{pi} } \OTarrow {\tau} \ostate{s^1_{h+1}\uplus s^{i\in L}_{pi}}}\]

\item For the observable open transition.  By Lemma \ref{lem-compose} we obtain:\\ 
	\[ P[Q]_{j_0}  
	\models
	{\openrule
		{
			{\beta^i_{p j}}^{(j\in J\setminus\{j_0\})} \uplus \set{\beta_2}, 
			\Pred_p\land\Pred^2\land \alpha_t=\beta^i_{p j_0 },  
			\Post_p\uplus \Post^2}
{\ostate{s^{i\in L}_{pi}\uplus s_0^2} \OTarrow {\alpha_p} \ostate{s^{i\in L}_{(p+1)i}\uplus s_1^2}}
	}
	\]

%
%\[P[Q]_{j_0}\!\models		
%\openrule
%    {
%       {\beta^i_{p j}}^{(j\in J\setminus\{j_0\})} \uplus \set{\beta_2},{\Pred_p}\!\!'',{\Post_p}\!\!''   }
%         {\ostate{s^{i\in L \uplus L_Q}_{pi}} \OTarrow {\alpha} \ostate{s^{i\in L\uplus L_Q }_{(p+1)i}}}\] \\
% such that  ${\Pred_p}\!\!'' \iff {\Pred_p}
%		\land {\Pred_p}\!\!'\land \alpha_t=\beta_{j_0}$, ${\Post_p}\!\!''={\Post_p}\uplus 
%		{\Post_p}\!\!'$\\ 
\item We proceed in the same way as the first item for $k'$ last weak open transitions, and we obtain $k'$ open tau transitions.\\
\end{enumerate}
